---
title: "P8105 Homework 3"
author: "sarah_8105"
output: github_document
---

This is my third homework assignment for P8105.

```{r libraries}
library(p8105.datasets)
```


## Problem 1

In this first problem, I explore the Instacart data. I create a plot of the number of items ordered in each aisle, limiting to aisles with more than 10,000 items ordered. 

```{r insta}
data("instacart")

```

Next, I create a table that displays the three most population items in each of the aisles of "baking ingredients", "dog food care", and "packaged vegetables fruits". The table displays the number of times each item was ordered.

```{r insta_t1}

```
In the final portion of problem 1, I create a table that shows the mean hour of the day at which Pink Lady apples and coffee ice cream are ordered on each day of the week.
```{r insta_t2}

```

The Instacart data contain de-identified grocery orders for users of this online grocery service, which partners with local stores like Whole Foods and Fairway to deliver groceries within 2 hours of order. Each row in the data set represents a product that has been ordered. The data include variables about the order (such as date and time), the customer, and about the products ordered (such as product name and department/aisle where the product can be found). 

There are a total of  `r ` orders, `r ` customers, and `r ` products represented in these data. On average, customers have placed `r ` orders and each order contains `r ` products. 

There are `r ` aisles represented in the data. Most items that are ordered come from the `r ` aisle, with `r ` items ordered. The most popular item from the baking ingredients aisle was `r `, the most population item from the dog food care aisle was `r `, and the most population item from the packaged vegetables fruits aisle was `r `.







 
## Problem 2

